### 2016-06-27 publications/Journal/3-features.tex
Updating section 3 complete.
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

A user can request a runtime switch between synchronization protocols. A typical use case is when a significant slowdown is observed or memory usage reaches a certain treshold. The simulation engine will swap out the current kernels with new instances while preserving the state of the models and the simulation itself, and resume using the new synchronization protocol. There is no real limit on the number of switches, but the replacing and synchronizing of the kernels comes at a runtime cost that is dependent on the number of present kernels and the actual hardware the simulation runs on. A test case requires for 2 kernels an overhead approaching a second. This is an experimental but promising feature that can be used in an automated framework driven by machine learning techniques.

@@

As we will demonstrate in ~\nameref{sec:4-subsec:overhead-pgraph:memory} memory allocation is a performance bottleneck in simulation with high event frequency, especially in parallel with optimistic synchronization. Dxex has a modular allocation interface that separates the object creation in kernels and models from the actual implementation. This allows us by using a single compilation parameter to change the allocation strategy. This interface allows instrumentation of memory usage. Not only useful to detect memory leaks should this be a concern, it can also be of use to a model author to trace event creation in large models. The default implementation is the fastest common configuration verified by the benchmarks used in this publication.

@@

A high performance asynchronous buffered threaded logging module is used in dxex which offers a complete overview of the entire simulation. The logging functionality can be used in any model, it only requires linking to the dxex library. Logging statements are written to file with a granular level system. In case of a severe crash the logging system will ensure persistence of the log file itself up until the last statement. Logging can be completely disabled to remove the performance impact.

@@

Invaluable during the development of dxex were runtime statistics collected by a simulation kernel. These record any measurable event in the simulation such as : events created/destroyed, inter kernel event frequency, reverts, GVT calculations, lookahead and eot values and insights into the fairness between the kernels themselves. Using this information it is possible to detect an imbalance between the simulation rounds that a single kernel executes compared to dependent or influencing kernels, observe how balanced inter kernel message traffic actually is during a simulation or measure the reverted simulation transition in optimistic synchronization.
The same information can, as in the visualization tool, be used programmatically to extend dxex with visualization software that shows the simulation from a white box perspective or allow activity tracking.
Figures \ref{fig:pholdtree_visualize_seq},\ref{fig:pholdtree_visualize_parDFS},\ref{fig:pholdtree_visualize_parBFS} give an example of the modular visualization output generated by dxex after simulation the PholdTree model. These figures demonstrate the effect of allocation visually, the high inter kernels event count will degrade performance, as we will demonstrate in 
\nameref{PholdTreeallocation} . Dotted lines represent connections between models that did not see events exchanged, directed edges represents events with frequency as weight and color to distinguish inter versus intra kernel events.
% Add figure
\begin{figure}
	\center
	\includegraphics[width=\plotfraction\columnwidth, height=6cm, keepaspectratio]{fig/pholdtreed1n3t5000.eps}
	\caption{Visualization of PholdTree (d=1,n=3,t=5000) sequential simulation.}
	\label{fig:pholdtree_visualize_seq}
\end{figure}
\begin{figure}
	\center
	\includegraphics[width=\plotfraction\columnwidth,  height=6cm, keepaspectratio]{fig/pholdtreed1n3t5000c4BFS.eps}
	\caption{Visualization of PholdTree (d=1,n=3,t=5000) parallel simulation with breadth first allocation and 4 kernels.}
	\label{fig:pholdtree_visualize_parBFS}
\end{figure}
\begin{figure}
	\center
	\includegraphics[width=\plotfraction\columnwidth,  height=6cm, keepaspectratio]{fig/pholdtreed1n3t5000c4DFS.eps}
	\caption{Visualization of PholdTree (d=1,n=3,t=5000) parallel simulation with depth first allocation and 4 kernels.}
	\label{fig:pholdtree_visualize_parDFS}
\end{figure}

>>>>

A user can request a runtime switch between synchronization protocols. A typical use case is when a significant slowdown is observed or memory usage reaches a certain treshold. The simulation engine will swap out the current kernels with new instances while preserving the state of the models and the simulation itself, and resume using the new synchronization protocol. There is no real limit on the number of switches, but the replacing and synchronizing of the kernels comes at a runtime cost that is dependent on the number of present kernels and the actual hardware the simulation runs on. In a test case with 2 kernels the overhead of a double runtime switch is less than a second.
This is an experimental but promising feature that can be used in an automated framework driven by machine learning techniques.

@@

As we will demonstrate in ~\nameref{sec:4-subsec:overhead-pgraph:memory} memory allocation is a performance bottleneck in simulations with high event frequency, especially in parallel with optimistic synchronization. Dxex has a modular allocation interface that separates the object creation in kernels and models from the actual implementation. This allows us by using a single compilation parameter to change the allocation strategy. This interface allows instrumentation of memory usage. Not only useful to detect memory leaks should this be a concern, it can also be of use to a model author to trace event creation in large models. The default implementation is the fastest common configuration verified by the benchmarks in this publication.

@@

A high performance asynchronous buffered logging module is used in dxex which offers a complete overview of the entire simulation. The logging functionality can be used in any model, it only requires linking to the dxex library. Logging statements are written to file with a granular level system. In case of a severe crash the logging system will ensure persistence of the log file itself up until the last statement. Logging can be completely disabled to remove the performance impact.

@@

Dxex optionally collects runtime statistics for each kernel. These record any measurable event in the simulation such as : events created/destroyed, inter/intra kernel event frequency, reverts, GVT calculations, lookahead and eot values and insights into the fairness between the kernels themselves. Using this information it is possible to detect an imbalance between the simulation rounds that a single kernel executes compared to dependent or influencing kernels, observe how balanced inter kernel message traffic actually is during a simulation or measure reverted transitions in optimistic synchronization.
The same information can, as in the visualization tool, be used programmatically to extend dxex with visualization software that shows the simulation from a white box perspective or provide feedback to drive synchronization switching.
Figures \ref{fig:pholdtree_visualize_seq},\ref{fig:pholdtree_visualize_parDFS},\ref{fig:pholdtree_visualize_parBFS} give an example of the modular visualization output generated by dxex after simulation the PholdTree model. These figures demonstrate the effect of allocation visually and show the high inter kernels event count that degrades performance, as we will demonstrate in 
\nameref{PholdTreeallocation} . Dotted lines represent connections between models where no events were exchanged, directed edges represents events with frequency as weight, coloring is used to distinguish inter versus intra kernel events.
% Add figure
\begin{figure}
	\center
	\includegraphics[width=\plotfraction\columnwidth, height=6cm, keepaspectratio]{fig/pholdtreed1n3t5000.eps}
	\caption{Visualization of PholdTree (d=1,n=3,t=5000) sequential simulation.}
	\label{fig:pholdtree_visualize_seq}
\end{figure}
\begin{figure}
	\center
	\includegraphics[width=\plotfraction\columnwidth,  height=6cm, keepaspectratio]{fig/pholdtreed1n3t5000c4BFS.eps}
	\caption{Visualization of PholdTree (d=1,n=3,t=5000) parallel simulation with breadth first allocation and 4 kernels.}
	\label{fig:pholdtree_visualize_parBFS}
\end{figure}
\begin{figure}
	\center
	\includegraphics[width=\plotfraction\columnwidth,  height=6cm, keepaspectratio]{fig/pholdtreed1n3t5000c4DFS.eps}
	\caption{Visualization of PholdTree (d=1,n=3,t=5000) parallel simulation with depth first allocation and 4 kernels.}
	\label{fig:pholdtree_visualize_parDFS}
\end{figure}
